Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-03-26 code-tests.tex
working on conduction test - anisotropic only.

\subsubsection{Representative thermal conduction test}
\label{sec.tests.conduct}
\red{(Brian)}
One test for the thermal conduction (maybe 1 for isotropic, 1 for
anisotropic).  Figure~\ref{fig.conduct}.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/iso_conduction_final_output.eps}
\includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
\caption{Temperature blah blah.  Conduction tests!}
\label{fig.conduct}
\end{center}
\end{figure}

>>>>

\subsubsection{Anisotropic Thermal conduction}
\label{sec.tests.conduct}

In a plasma, conduction takes place primarily due to electron motion.
In astrophysical environments such as the intergalactic and
intracluster media, with typical magnetic strengths of $\sim 1 \mu$G
and temperatures in the millions of Kelvin, the electron gyroradius is
very small compared to the scapes of interest in a cosmological
simulation.  As a result, electrons can only move \textit{along}
magnetic field lines, not perpendicular to them, which can result in
very interesting magnetohydrodynamical instabilities
\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}. 
 The correct modeling of this behavior can be quite challenging, and is described in 
Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is calculated
in both an operator-split and directionally-split manner, and the
addition of heat transport solely along magnetic field lines requires
computation of cross-terms in the temperature derivative at cell
faces, which can add spurious oscillations in the temperature field in
regions where the temperature gradient is strong in more than one
spatial direction unless an appropriate flux-limiter \citep[such as that
of][]{1977JCoPh..23..263V} is chosen.

Figure~\ref{fig.conduct} shows a test that demonstrates the
correct behavior of anisotropic thermal conduction in \enzo.  We
initialize a two-dimensional, $256^2$ cell simulation having a physical
scale of 1 kpc on a side, a uniform density of 1 proton/cc, and a
background temperature of $10^6$ Kelvin.  Magnetic fields with a
strength of  B$_0 = 1 \mu$G are initialized such that the field lines form
circles around the center of the simulation volume, such that B$_x =
-B_0\sin(\theta)$ and B$_y = B_0\cos(\theta)$, where $\theta$ is the
angle measured from the $+x$ direction in a counterclockwise manner.
A Gaussian temperature pulse is injected at $(0.75, 0.5)$ (in units of the box
size), with a peak temperature of $10^8$ K and a FWHM of $0.015625$ of
the box size.  This initial setup is shown in the left panel of
Figure~\ref{fig.conduct}.  The simulation is then allowed to evolve
with \textit{only} anisotropic conduction turned on (e.g.,
no hydrodynamics or radiative cooling), and with a Spitzer fraction of
f$_{sp} = 1$.

The right panel of Figure~\ref{fig.conduct} shows the state of the
simulation after 300 Myr.  Heat has clearly been transported only along
field lines -- there has been no diffusion perpendicular to the
magnetic field setup, which is critical for many studies involving
anisotropic thermal conduction.  No oscillations are seen in the
temperature field in regions where the fields are not aligned with the
grid, suggesting that the flux-limiter is operating as expected.

\begin{figure}
\begin{center}
\includegraphics[width=0.42\textwidth]{figures/aniso_conduction_initial_output.eps}
\includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
\caption{Two-dimensional anisotropic conduction test in a uniform,
constant temperature background with circular magnetic fields centered
on (0.5, 0.5).  The background medium has a density of 1 proton/cc and
a temperature of $10^6$~K.  At t$ = 0$ (left panel), a Gaussian heat
pulse is injected at (0.75, 0.5) with a FWHM of $0.015625$ (with all
numbers given in units of the box size) and a peak temperature of
$10^8$~K, and allowed to evolve without hydrodynamical motion (i.e.,
static gas) and no radiative cooling for 300 Myr.  At t$ = 300$~Myr
(right panel), heat has been transported along magnetic field lines
with no significant diffusion perpendicular to field
lines. Furthermore, there is no detectable oscillations in the
temperature in regions where the field is not parallel with the grids.}
\label{fig.conduct}
\end{center}
\end{figure}
